Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-05-30 removals/manuscript/removal_ms.tex
ms edits by SPE

We simulated the IPM to estimate equilibrium abundances under three different scenarios. For all IPM simulations, we set the spatial random effects to zero, meaning that we are simulating dynamics on a hypothetical, average site. The first scenario featured all four species and our baseline model for control plots (no removal effects). We initialized all species with arbitrarily low cover and ran the model for 500 time steps (years) to allow the community to reach an equilibrium. To incorporate temporal environmental variation, at each time step we randomly selected one of the 30 sets of random year effects for the survival, growth and recruitment models (``kernel selection'' in the terminology of \citealt{metcalf_statistical_2015}). We calculated the equilibrium cover for each species by simulating an additional 2000 time steps and averaging each species' cover over this period. For the second scenario, we set the cover of \textit{A. tripartita} to zero and then repeated the procedure. This simulation projects the competitive release of grasses following \textit{A. tripartita} removal according to the baseline model (no removal effects). For the third scenario, we again set \textit{A. triparitia} cover to zero, but this time we projected the model using the vital rate regressions that include removal effects. If our baseline model successfully predicts competitive release observed in the removal experiment, then the results of the second and third scenarios should be similar. To the extent that the baseline model underestimates (or overestimates) the actual competitive release, the results of the second and third scenarios will differ. As with the IBM, we repeated these simulations using maximum removal treatment effects represented by parameters values at the 5\% or 95\% confidence limit farthest from zero. We do not show simulations of long-term \textit{A. tripartita} response to grass removal because the removal treatment effects in the vital rate regressions were never significantly different from zero, indicating that the baseline model effectively predicted competitive release. \comment{PBA: Just do it? SPE: Just do it, and then explain why there is no difference between ``grass removal'' and ``grass removal with removal treatment effect''. }

@@

 The equilibrium cover of \textit{A. tripartita} simulated by the IPM (6.0\%) is consistent with observed cover during the historical data collection period (6.6\% from 1931-1956), but is low compared to contemporary observed cover (15-25\%, Fig. \ref{fig:CoverTrends}). Could the small competitive release following simulated \textit{A. tripartita} removal simply reflect the low \textit{A. tripartita} cover? To explore this possibility, we increased \textit{A. tripartita} cover in the baseline simulation by increasing average \textit{A. tripartita} fecundity. The increased fecundity led to a doubling of \textit{A. tripartita} average cover, but had essentially zero effect on the cover of the grasses (changes of less than 0.1\% cover).

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate 
the intensity of interspecific competition. After accounting 
for local neighborhood crowding, the removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; 
our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to removal.  
\textit{P. secunda} showed mixed responses to \textit{A. tripartita} removal: removal effects were not significant at the survival stage, 
were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{A. tripartita} 
on \textit{P. secunda} is positive). At the population level, models with and without removal effects projected similar patterns of 
short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata} and \textit{A. tripartita}. 
For \textit{P. secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps 
because the short-term projections mostly reflect the positive effect of \textit{A. tripartita} removal on growth, while the long-term 
equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did 
not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level 
consequences of removals on these three species.

>>>>

We simulated the IPM to estimate equilibrium abundances under three different scenarios. For all IPM simulations, we set the spatial random effects to zero, meaning that we are simulating dynamics on a hypothetical, average site. The first scenario featured all four species and our baseline model for control plots (no removal effects). We initialized all species with arbitrarily low cover and ran the model for 500 time steps (years) to allow the community to reach an equilibrium. To incorporate temporal environmental variation, at each time step we randomly selected one of the 30 sets of random year effects for the survival, growth and recruitment models (``kernel selection'' in the terminology of \citealt{metcalf_statistical_2015}). We calculated the equilibrium cover for each species by simulating an additional 2000 time steps and averaging each species' cover over this period. For the second scenario, we set the cover of \textit{A. tripartita} to zero and then repeated the procedure. This simulation projects the competitive release of grasses following \textit{A. tripartita} removal according to the baseline model (no removal effects). For the third scenario, we again set \textit{A. triparitia} cover to zero, but this time we projected the model using the vital rate regressions that include removal effects. If our baseline model successfully predicts competitive release observed in the removal experiment, then the results of the second and third scenarios should be similar. To the extent that the baseline model underestimates (or overestimates) the actual competitive release, the results of the second and third scenarios will differ. As with the IBM, we repeated these simulations using maximum removal treatment effects represented by parameters values at the 5\% or 95\% confidence limit farthest from zero. We do not show simulations of long-term \textit{A. tripartita} response to grass removal because the removal treatment effects in the vital rate regressions were never significantly different from zero, indicating that the baseline model effectively predicted competitive release.

@@

The equilibrium cover of \textit{A. tripartita} simulated by the IPM (6.0\%) is consistent with observed cover during the historical data collection period (6.6\% from 1931-1956), but is low compared to contemporary observed cover (15-25\%, Fig. \ref{fig:CoverTrends}). Could the small competitive release \new{of the grasses} 
following simulated \textit{A. tripartita} removal simply reflect the low simulated \textit{A. tripartita} cover? To explore this possibility, we increased \textit{A. tripartita} cover in the baseline simulation by increasing its average fecundity. The increased fecundity led to a doubling of \textit{A. tripartita} average cover, but had essentially zero effect on the cover of the grasses (changes of less than 0.1\% cover), \new{implying that shrub removal from higher cover 
would still have little effect on the grasses.}

\new{The baseline model predicts more dramatic competitive release of \textit{A. tripartita} following grass removal: mean cover increases from 5.6\% to 16.2\%. Adding
removal treatment effects dramatically alters the outcome, with \textit{A. tripartita} decreasing towards extinction. However, these outcomes both are driven
by parameters that are estimated with very high uncertainty (Table~\ref{table:recruitment}). The baseline model predicts an increase following grass removal because of the relatively strong negative effects of \textit{Poa secunda} and \textit{H. comata} on \textit{A. tripartita} recruitment (\ref{table:recruitment}). The estimated effect of
\emph{Poa} is -0.66, but the standard deviation of the estimate is 0.33 and the 95\% credible interval stops just short of zero; the 95\% credible interval for the
\textit{H. comata} effect includes values less than half of the point-estimate. The grass removal treatment effect, which leads to predicted extinction, has a credible 
interval that broadly overlaps zero (Table~\ref{table:recruitment}). \textit{A. tripartita} recruitment parameters have exceptionally high uncertainty, presumably 
because \textit{A. tripartita} recruitment is episodic (see \emph{Discussion}), and this produces extreme uncertainty about its long-term 
response to grass removal.}

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate 
the intensity of interspecific competition. After accounting 
for local neighborhood crowding, the removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; 
our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to removal.  
\textit{P. secunda} showed mixed responses to \textit{A. tripartita} removal: removal effects were not significant at the survival stage, 
were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{A. tripartita} 
on \textit{P. secunda} is positive). At the population level, models with and without removal effects projected similar patterns of 
short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata}. \new{ For \textit{A. tripartita}, 
the experimental observations and short-term population projections show little response to removal, while the long-term simulations 
from the baseline and removal effects models suggest strong (but opposite) effects of removal: a tripling of abundance in the baseline model, and extinction in the
removal effects model. However, these long-term simulations primarily reflect the influence of recruitment parameters 
that are highly uncertain (see \emph{Results}). This high uncertainty 
results from \textit{A. tripartita} recruitment being low (a mean of 0.63 recruits per quadrat per year, compared to means of 2.6 to 5.3 for the grasses) and 
very episodic (roughly half of the total recruits in the 22 years of historical data were observed in 2 of those years, and only 13 \textit{A. tripartita} 
recruits were observed since we resumed annual censuses in 2007). Given this uncertainty, we put more weight on the direct observations 
and short-term projections showing little effect of removal on \textit{A. tripartita} populations.}
For \textit{P. secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps 
because the short-term projections mostly reflect the positive effect of \textit{A. tripartita} removal on growth, while the long-term 
equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did 
not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level 
consequences of removals on these three species.
